\documentclass[11pt]{article}
\usepackage{amsmath,amssymb,amsthm,stmaryrd}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\title{Grieg: A Four-Fold Phase Semantics on an Ontic Logical Manifold}
\author{William A. Patterson}
\date{\today}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{definition}{Definition}

\begin{document}
\maketitle

\begin{abstract}
Grieg augments classical truth with an ontic phase over $\{\mathsf{ALIVE},\mathsf{JAM},\mathsf{MEM},\mathsf{VAC}\}$ capturing boundary, transport, and witness on a logical manifold. We give a sequent calculus, prove conservativity over classical propositional logic on phase-free fragments, show an embedding of Belnap--Dunn with a separation (no phase-erasing compositional translation), establish NP/coNP bounds, and present a trace semantics whose invariants (sink on $\to$, $\pi$ rotation on $\neg$, etc.) are observational and do not alter results.
\end{abstract}

\section{Syntax}
Atoms $p\in\mathcal{P}$; connectives $\neg,\wedge,\vee,\to$; phase operators $\mathsf{@mem}(\cdot)$, $\mathsf{@jam}(\cdot)$, $\mathsf{@vac}(x)$, $\mathsf{@alive}(\cdot)$.

\section{Semantics}
Truth values $V=\{0,1,\bot\}$; phases $P=\{\mathsf{ALIVE},\mathsf{JAM},\mathsf{MEM},\mathsf{VAC}\}$. A model $\mathcal{M}=\langle W,\nu,\phi\rangle$ with boolean evaluation $\llbracket\cdot\rrbracket_w:\mathrm{Form}\to V$ (classical on defined points) and phase map $\phi:\mathrm{Form}\times W\to P$ satisfying invariants: negation rotates on the factual sheet, disjunction is centrifugal with boundary short-circuit, implication has a monotone radial sink under modus ponens, $\mathsf{@mem}$ transports between sheets, $\mathsf{@vac}$ projects to counterfactual where witness is absent.


\subsection{Sheets (F:C), Transport, and Sinks}
Each sheet is a solid torus $\Sigma = S^1 \times D^2$ with angles $(\theta,\varphi)$ and radial potential $\rho \in [0,1]$.
We evaluate on two ontic sheets: $\mathbf{F}$ (factual) and $\mathbf{C}$ (counterfactual).

Judgments:
\[
  \langle s,\phi \rangle \;\vdash\; E \Downarrow (v,\phi',s')
\]
with $s \in \{\mathbf{F},\mathbf{C}\}$, $\phi \in \{\mathsf{ALIVE},\mathsf{JAM},\mathsf{MEM},\mathsf{VAC}\}$,
and $v \in \{\mathsf{true},\mathsf{false},\bot\}$ ($\bot =$ no total boolean).

\paragraph{VAC projection.}
\[
\frac{}{ \langle s,\phi \rangle \vdash @\!vac(x) \Downarrow (\bot,\mathsf{VAC},\mathbf{C}) }.
\]

\paragraph{MEM transport (truth-preserving).}
Let $\overline{s}$ be the opposite sheet. Then
\[
\frac{ \langle \overline{s},\mathsf{MEM} \rangle \vdash E \Downarrow (v,\phi_1,\overline{s}) }
     { \langle s,\phi \rangle \vdash @\!mem(E) \Downarrow (v,\mathsf{MEM},s) }.
\]

\paragraph{Implication and F-local sinks.}
On $\mathbf{F}$, if the antecedent is established, implication chains flow toward the core ($\rho \downarrow 0$) and may mark a \emph{sink} (absorbing point).
On $\mathbf{C}$, sinks are never registered.

\paragraph{Conservativity.}
Restricted to $\mathbf{F}$ without phase operators, the boolean projection agrees with CPL; sheet/phase are observational structure.

\section{Proof System $\mathbf{G\!r}$}
Classical propositional rules; admissible phase rules enforcing invariants without changing boolean projection.

\begin{theorem}[Conservativity over CPL]
For every phase-free formula $F$, $F$ is valid in Grieg iff $F$ is valid in classical propositional logic.
\end{theorem}
\begin{proof}[Sketch]
Define projection $\pi:(v,p)\mapsto v$. By structural induction, truth-clauses coincide with CPL and no phase operator is present to produce non-$\mathsf{ALIVE}$ on closed formulas; hence $\pi(\llbracket F\rrbracket)=\llbracket F\rrbracket_{\mathrm{CPL}}$.
\end{proof}

\begin{theorem}[Embedding + Separation vs Belnap--Dunn]
There exists a translation $E:\mathsf{FOUR}\to$ Grieg preserving boolean projection. However, there is no compositional translation $T$ from Grieg to $\mathsf{FOUR}$ that preserves satisfaction and context-equivalence while erasing phases.
\end{theorem}
\begin{proof}[Idea]
Map BD's `both/neither` strata into $\mathsf{JAM}/\mathsf{VAC}$ with matching boolean projection. For separation, build contexts $C[\_]$ where Grieg distinguishes formulas that differ only by an \emph{operational sink} (licensed modus ponens) while BD assigns the same four-value because it is extensional and phase-blind. Compositionality of $T$ would force $C[T(F_1)]=C[T(F_2)]$, contradicting Grieg's distinction.
\end{proof}

\section{Trace Semantics}
Define per-step records with $(\mathrm{op},\mathrm{pre},\mathrm{post},\mathrm{sink},\mathrm{jam},\theta,\rho)$. Prove observational adequacy (I1–I5) by induction on evaluation.

\section{Complexity}
SAT(Grieg) is NP-complete and VAL(Grieg) is coNP-complete by reduction from/to CPL and polynomial verification of local phase constraints.

\section{Implementation \& Tests}
Reference Rust engine; property tests for I1–I5; traces are feature-gated and semantics-preserving.

\bibliographystyle{plain}
\bibliography{grieg}
\end{document}
